---
title: "Prediciton Project 1"
author: "Stella Veazey"
date: "4/20/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(Hmisc)
library(caret)
library(pROC)
thorsurg <- read.csv("~/Desktop/Predictive/project 1/ThoraricSurgery.arff.txt", header=FALSE)
names(thorsurg) <- c("DGN", "FVC", "FEV1", "PRZ", "pre.pain", "pre.haem", "pre.dysp", "pre.cough", "pre.weak", "TNM", "T2DM", "MI6", "PAD", "smoke", "asthma", "AGE", "RISK1YR")

describe(thorsurg)

#near zero variance predictors
nzv <- nearZeroVar(thorsurg, saveMetrics = TRUE)
nzv[nzv$nzv,][1:3,]

#mycardial infarction lasrt 6 months
table(thorsurg$MI6==1,thorsurg$RISK1YR)

#asthma
table(thorsurg$asthma==1, thorsurg$RISK1YR)

#peripheral arterial diseases
table(thorsurg$PAD==1, thorsurg$RISK1YR)
```
Notes:

*DGN1, DGN5, DGN4, DGN8 are sparse

*PRZ 2 only 27 (Performance status - Zubrod scale)

*pre.pain true = 27

*pre.haem true = 68

*pre.dysp = true for 31

*TNM (tumor size) oc13 = 19 and oc14=17 (largest sizes)

*T2DM (diabetes) true=35

*MI6 (heart) true=2

*PAD true=8

*asthma true = 2

*Risk1yr true=70

```{r}
#recoding
thorsurg$RISK1YR <- ifelse(thorsurg$RISK1YR=="TRUE", 1, 0)
thorsurg$asthma <- ifelse(thorsurg$asthma=="TRUE", 1, 0)
thorsurg$smoke <- ifelse(thorsurg$smoke=="TRUE", 1, 0)
thorsurg$PAD <- ifelse(thorsurg$PAD=="TRUE", 1, 0)
thorsurg$MI6 <- ifelse(thorsurg$MI6=="TRUE", 1, 0)
thorsurg$T2DM <- ifelse(thorsurg$T2DM=="TRUE", 1, 0)
thorsurg$pre.weak <- ifelse(thorsurg$pre.weak=="TRUE", 1, 0)
thorsurg$pre.cough <- ifelse(thorsurg$pre.cough=="TRUE", 1, 0)
thorsurg$pre.dysp <- ifelse(thorsurg$pre.dysp=="TRUE", 1, 0)
thorsurg$pre.haem <- ifelse(thorsurg$pre.haem=="TRUE", 1, 0)
thorsurg$pre.pain <- ifelse(thorsurg$pre.pain=="TRUE", 1, 0)

thorsurg$DGN <- ifelse(thorsurg$DGN=="DGN1", 1, ifelse(thorsurg$DGN=="DGN2", 2, ifelse(thorsurg$DGN=="DGN3", 3, ifelse(thorsurg$DGN=="DGN4", 4, ifelse(thorsurg$DGN=="DGN5", 5, ifelse(thorsurg$DGN=="DGN6", 6, ifelse(thorsurg$DGN=="DGN7", 7, 8)))))))

thorsurg$PRZ <- ifelse(thorsurg$PRZ=="PRZ0", 0, ifelse(thorsurg$PRZ=="PRZ1", 1, 2))

thorsurg$TNM <- ifelse(thorsurg$TNM=="OC11", 1, ifelse(thorsurg$TNM=="OC12", 2, ifelse(thorsurg$DGN=="OC13", 3, 4)))

feature.names=names(thorsurg)
for (f in feature.names) {
  if (class(thorsurg[[f]])=="factor") {
    levels <- unique(c(thorsurg[[f]]))
    train[[f]] <- factor(thorsurg[[f]],
                   labels=make.names(levels))
  }
}

#caret was being weird about the levels of the output variable not being appropriate variable names when I was playing with a tree model, hence this:
thorsurg$RISK1YR <- ifelse(thorsurg$RISK1YR==1, "died", "lived")
```

LASSO (smote-free) 
AUC: 0.729
Youden threshold: 0.130
Specificity: 0.555
Sensitivity: 0.786

```{r}
library(glmnet)
library(Matrix)

#dropped MI6 and asthma because their very low variance would not allow model to run
xvars <-thorsurg[,-c(12,15,17)] 

x <- as.matrix(xvars)
y <- thorsurg$RISK1YR

ctrl <- trainControl(method="repeatedcv",   # 10fold cross validation
                     repeats=5,		    # do 5 repititions of cv
                     summaryFunction=twoClassSummary,	# Use AUC to pick the bestmodel
                     classProbs=TRUE)

#lasso
set.seed(1953)
lasso1 <- train(x=xvars, y=thorsurg$RISK1YR, method="glmnet", trControl=ctrl, metric="ROC")

#variable importance
importance <- varImp(lasso1, scale = FALSE)
plot(importance)

#ROC
pred.l1 <- predict(lasso1, type="prob")
risk.num <- as.numeric(ifelse(thorsurg$RISK1YR=="died", 1, 0))
summary(pred.l1)
roc.l1 <- roc(risk.num, pred.l1[,1])

#Youden's
youd.l1 <- coords(roc.l1, x = "best", best.method = "youden", ret = c("threshold", "specificity", "sensitivity", "accuracy", "npv", "ppv", "fn", "fp"))
round(youd.l1,2)
yl1<-round(youd.l1[1],2)


#plot
plot(roc.l1,print.thres=yl1, print.auc=TRUE)
```

Lasso with SMOTE

AUC: 0.719

Sensitivity: 0.729

Specificity: 0.613
```{r}
ctrl.smote <- trainControl(method="repeatedcv", 
                     number=10,
                     repeats=3,
                     classProbs=T,
                     summaryFunction=twoClassSummary,
                     sampling="smote")

set.seed(27)
lasso2 <- train(x=xvars, y=thorsurg$RISK1YR, method="glmnet", trControl=ctrl.smote, metric="ROC")

#variable importance
importance <- varImp(lasso2, scale = FALSE)
plot(importance)

#ROC
pred.l2 <- predict(lasso2, type="prob")
roc.l2 <- roc(risk.num, pred.l2[,1])

#Youden's
youd.l2 <- coords(roc.l2, x = "best", best.method = "youden", ret = c("threshold", "specificity", "sensitivity", "accuracy", "npv", "ppv", "fn", "fp"))
round(youd.l2,2)
yl2<-round(youd.l2[1],2)

#plot
plot(roc.l2,print.thres=yl2,print.auc=TRUE)

#confusion matrices
library(SDMTools)
confusion.matrix(risk.num, pred.l1[,1], threshold=yl1)
confusion.matrix(risk.num, pred.l2[,1], threshold=yl2)
```


Random Forest 
```{r, eval=FALSE, echo=FALSE}
library(randomForest)

ctrl <- trainControl(method="repeatedcv",   # 10fold cross validation
                     repeats=5,		    # do 5 repititions of cv
                     summaryFunction=twoClassSummary,	# Use AUC to pick the bestmodel
                     classProbs=TRUE)

rf1 <- train(x=thorsurg[,-17], y=factor(thorsurg$RISK1YR), method="rf", trControl=ctrl)

rf1.pred <- predict(rf1, thorsurg, type="prob")
summary(rf1.pred)

t1prb <- predict(tree1$finalModel, thorsurg)
plot(tree2$finalModel)
text(tree2$finalModel)

library(pROC)
auc <- roc(t2pred, thorsurg, type="class")
```

SVM 

AUC: 0.988

Sensitivity: 0.829

Specificity: 0.99

```{r}
#code from https://www.r-bloggers.com/the-5th-tribe-support-vector-machines-and-caret/
library(dplyr)     
library(kernlab)       

set.seed(1492)

#SVM
svm.tune1 <- train(x=xvars,
                  y= thorsurg$RISK1YR,
                  method = "svmRadial",   # Radial kernel
                  tuneLength = 9,					# 9 values of the cost function
                  preProc = c("center","scale"),  # Center and scale data
                  metric="ROC",
                  trControl=ctrl)

svm.tune1

set.seed(1492)
# Use the expand.grid to specify the search space	
grid <- expand.grid(sigma = c(.04, .06, 0.08, .1, .12),
                    C = c(0.4, .45, .5, .55, .6))

svm.tune2 <- train(x=xvars,
                    y= thorsurg$RISK1YR,
                    method = "svmRadial",
                    preProc = c("center","scale"),
                    metric="ROC",
                    tuneGrid = grid,
                    trControl=ctrl)

svm.tune2

#ROC
svm.pred2 <- predict(svm.tune2, type="prob")
roc.svm2 <- roc(risk.num, svm.pred2[,1])

#Youden's

youd.svm2 <- coords(roc.svm2, x = "best", best.method = "youden", ret = c("threshold", "specificity", "sensitivity", "accuracy", "npv", "ppv", "fn", "fp"))
round(youd.svm2,2)
ysvm <- round(youd.svm2[1],2)

#plot
plot(roc.svm2, print.thres=ysvm, print.auc=TRUE)
```


SVM using SMOTE

AUC: 0.905

Sensitivity: 0.914

Specificity: 0.915

```{r}
set.seed(456)
svm.tune3 <- train(x=xvars,
                  y= as.factor(thorsurg$RISK1YR),
                  method = "svmRadial",   # Radial kernel
                  tuneLength = 9,					# 9 values of the cost function
                  preProc = c("center","scale"),  # Center and scale data
                  metric="ROC",
                  trControl=ctrl.smote)

#ROC
svm.pred3 <- predict(svm.tune3,type="prob")
roc.svm3 <- roc(risk.num, svm.pred3[,1])

#Youden's
youd.svm3 <- coords(roc.svm3, x = "best", best.method = "youden", ret = c("threshold", "specificity", "sensitivity", "accuracy", "npv", "ppv", "fn", "fp"))
round(youd.svm3,2)
ysvm3 <- round(youd.svm3[1],2)

#plot
plot(roc.svm3,print.thres=ysvm3,print.auc=TRUE)

#confusion matrices
confusion.matrix(risk.num, svm.pred2[,1], threshold=ysvm)
confusion.matrix(risk.num, svm.pred3[,1], threshold=ysvm3)
```

GBM (smote-free)

AUC: 0.770

Sens: 0.829

Spec: 0.665
```{r}
set.seed(56)
gbm1 <- train(x=xvars, y=thorsurg$RISK1YR, method="gbm", distribution="bernoulli", metric="ROC", trControl=ctrl, verbose=FALSE)

pred.gbm1 <- predict(gbm1, type="prob")
roc.gbm1 <- roc(risk.num, pred.gbm1[,1])

youd.gbm1 <- coords(roc.gbm1, x = "best", 
                   best.method = "youden", 
                   ret = c("threshold", "specificity", 
                           "sensitivity")) 

round(youd.gbm1,2)
ygbm1 <- round(youd.gbm1[1],2)

#plot
plot(roc.gbm1,print.thres=ygbm1,print.auc=TRUE)
```

GBM with SMOTE

AUC: 0.721

Sens: 0.700

Spec: 0.693
```{r}
set.seed(565)
gbm2 <- train(x=xvars, y=thorsurg$RISK1YR, method="gbm", distribution="bernoulli", metric="ROC", trControl=ctrl.smote, verbose=FALSE)

pred.gbm2 <- predict(gbm2, type="prob")
roc.gbm2 <- roc(risk.num, pred.gbm2[,1])

youd.gbm2 <- coords(roc.gbm2, x = "best", 
                   best.method = "youden", 
                   ret = c("threshold", "specificity", 
                           "sensitivity")) 

round(youd.gbm2,2)
ygbm2 <- round(youd.gbm2[1],2)

#plot
plot(roc.gbm2,print.thres=ygbm2,print.auc=TRUE)

#confusion matrix
confusion.matrix(risk.num, pred.gbm1[,1], threshold=ygbm1)
confusion.matrix(risk.num, pred.gbm2[,1], threshold=ygbm2)
```